
# Results



## Descriptive statistics

```{r}
# Calculate number of workers
n <- length(unique(ab_panel$oracle_number[ab_panel$date >= '2012-01-01' & ab_panel$date <= '2016-12-31']))
```

We collected absenteeism and demographic data data on `r n` workers from 2013 through 2016. 

```{r, results='asis'}
# x <- ab_panel %>%
#   left_join(workers %>%
#               dplyr::select(oracle_number, date_of_birth, permanent_or_temporary, sex))
x <- model_data
x$age <-as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25
x$agey <- x$age
x$age <- round(x$age, digits = -1)
x$status <- x$permanent_or_temporary


calculate <- function(var= 'sex',
                      data){
  out <- 
    data %>%
    group_by_(var) %>%
    summarise(`Treatment workers` = length(unique(oracle_number[!months_since %in% c('Never', 'Before')])),
              `Treatment days` = length(oracle_number[!months_since %in% c('Never', 'Before')]),
              `Control workers` = length(unique(oracle_number[months_since %in% c('Never', 'Before')])),
              `Control days` = length(oracle_number[months_since %in% c('Never', 'Before')])) %>%
    ungroup() %>%
    # mutate(p = round(n / sum(n) * 100, digits = 2)) %>%
    mutate(variable = '') 
    names(out)[1] <- ' '
    out$variable[1] <- Hmisc::capitalize(var)
    
    out <- out %>% dplyr::select(variable, ` `, `Treatment workers`, `Treatment days`, `Control workers`, `Control days`)
    out$` ` <- as.character(out$` `)
  return(out)
}
out_list <- list()
vars <- c('department', 
           'sex',
           'status',
           'age')
for(i in 1:length(vars)){
  message(i)
  y <- calculate(var = vars[i], data = x)
  out_list[[i]] <- y
}
demo_table <- bind_rows(out_list)
names(demo_table)[1:2] <- c('Variable', ' ')
kable(demo_table, format = 'html') %>% kable_styling()
# print(xtable::xtable(demo_table,
#                      caption = 'Overall worker characteristics',
#                      align = c('r', 'r', rep('l', 5))),
#       include.rownames = FALSE)

```


```{r, fig.align='center', fig.cap = 'Age distribution by sex and worker status'}
# x$age <- round(as.numeric(as.Date('2014-01-01') - x$date_of_birth) / 365.25, digits = 2)
g1 <- ggplot(data = x,
       aes(x = jitter(agey, factor = 1000),
           group = sex,
           fill = sex)) +
  ggridges::geom_density_ridges(aes(y = sex,
                                    fill = sex),
                                height = 0.2) +
  facet_wrap(~permanent_or_temporary) +
  theme_maragra() +
  labs(x = 'Age',
       y = 'Density') +
  scale_fill_manual(name = 'Sex',
                    values = c('lightblue', 'grey'))
g1
```


```{r, fig.width = 6, fig.height = 3.5, fig.align='center', fig.cap = 'i. Monthly total precipitation in the Maniça district; ii. Average monthly temperature (bars) during the study period, as well as monthly maximum (red) and minimum (blue) temperatures'}
x <- weather %>%
  mutate(month = format(date, '%m')) %>%
  mutate(year = format(date, '%Y')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date) %>%
  summarise(precipitation = sum(precipitation, na.rm = TRUE),
            temp_max = max(temp_max, na.rm = TRUE),
            temp_min = min(temp_min, na.rm = TRUE),
            temp = mean(temp))
# x <- gather(x, key, value,precipitation:temp)

g1 <- ggplot(data = x,
       aes(x = date,
           y = precipitation)) +
  geom_bar(stat = 'identity',
           fill = 'lightblue',
           size = 0.1) +
  theme_maragra() +
  theme(axis.text.x = element_text(angle = 0)) +
  labs(x = 'Quarter',
       y = 'Milimeters') +
  scale_x_date(labels = scales::date_format("%Y"))

g2 <- ggplot(data = x,
            aes(x = date,
                y = temp)) +
  geom_bar(stat = 'identity', fill = 'grey', size = 0.1) +
  geom_line(aes(y = temp_max),
             color = 'darkred', 
             alpha = 0.5) +
  geom_line(aes(y = temp_min),
             color = 'blue', 
             alpha = 0.5) +
  # geom_linerange(aes(ymin = temp_min,
  #                    ymax = temp_max),
  #                alpha = 0.6,
  #                color = 'blue') +
   theme_maragra() +
  theme(axis.text.x = element_text(angle = 0)) +
  labs(x = 'Quarter',
       y = 'Degrees (Celcius)') +
  scale_x_date(labels = scales::date_format("%Y"))
Rmisc::multiplot(g1, g2, cols = 2)
```



```{r, fig.cap = 'Clinical malaria (district of Manhiça), all-cause absenteeism among Maragra workers, sick absenteeism among Maragra workers, positive cases at company clinic, and test positivity rate at company clinic', fig.height=5, fig.width=8}
# BES incidence
g1 <- 
  ggplot(data = bes %>%
           mutate(season = ifelse(p >= median(p), 'High season',
                                  'Low season')) %>%
           filter(weekdays(date) == 'Monday'),
       aes(x = date,
           y = p)) +
  geom_point(alpha = 0.6,
             aes(color = season),
             size = 0.2) +
  geom_line(alpha = 0.1) +
  theme_maragra(base_size = 9) +
  labs(x = 'Date',
       y = 'Incidence',
       title = 'A: Annualized district clinical\nmalaria incidence per 1,000') +
  scale_color_manual(name = '',
                     values = c('red', 'black')) + 
  guides(color = guide_legend(ncol = 1)) +
  theme(legend.position = c(.8, .8),
        legend.text = element_text(size = 5),
        legend.background = element_rect(fill=alpha('white', 0)))

# Absenteeism rate
g2_data <- ab_panel %>%
  group_by(month = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
  summarise(absences = length(which(absent)),
            absences_sick = length(which(absent_sick)),
            
            eligibles = n()) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100,
         sick_absenteeism_rate = absences_sick / eligibles * 100)

g2 <- ggplot(data = g2_data,
             aes(x = month,
                 y = absenteeism_rate)) +
  geom_line(alpha = 0.6) +
  theme_maragra(base_size = 9) +
  labs(x = 'Date',
       y = 'Absenteeism rate',
       title = 'B: Monthly absenteeism rate')

g3 <- ggplot(data = g2_data,
             aes(x = month,
                 y = sick_absenteeism_rate)) +
  geom_line(alpha = 0.6) +
  theme_maragra(base_size = 9) +
  labs(x = 'Date',
       y = 'Absenteeism rate',
       title = 'C: Monthly sick absenteeism rate') 

# Weather
g4 <-
  ggplot(data = weather %>%
           group_by(date = as.Date(paste0(format(date, '%Y-%m'), '-01'))) %>%
           summarise(precipitation = sum(precipitation, na.rm = TRUE)),
         aes(x = date,
             y = precipitation)) +
  geom_line(alpha = 0.6) +
  theme_maragra(base_size = 9) +
  labs(x = 'Date',
       y = 'Mililiters',
       title = 'D: Monthly precipitation') 

# Clinical incidence of malaria
g5 <- clinic_agg %>%
  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date) %>%
  summarise(tested = sum(tested),
            positive = sum(positive)) %>%
  ungroup %>%
  mutate(percent_positive = positive / tested * 100) %>%
  ggplot(aes(x = date,
             y = positive)) +
  geom_line(alpha = 0.6) +
  theme_maragra(base_size = 9) +
  labs(x = 'Date',
       y = 'Cases',
       title = 'D: Monthly malaria cases\nat company clinic') 

g6 <- clinic_agg %>%
  filter(date >= '2013-01-01', date <= '2016-12-31') %>%
  # mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date) %>%
  summarise(tested = sum(tested),
            positive = sum(positive)) %>%
  ungroup %>%
  mutate(percent_positive = positive / tested * 100) %>%
  ggplot(aes(x = date,
             y = percent_positive)) +
  geom_line(alpha = 0.6) +
  theme_maragra(base_size = 9) +
  labs(x = 'Date',
       y = 'Cases',
       title = 'E: Monthly malaria test positivity rate\nat company clinic') 
Rmisc::multiplot(g1, g2, g3,
                 # g4, 
                 g5, g6, layout = matrix(c(1, 1, 2, 3, 4, 5), nrow=2, byrow=TRUE)) 
```


```{r}
fumigations <- mc %>%
  filter(date >= '2013-01-01',
         date <= '2016-12-31')
n_fumigations <- nrow(fumigations)
n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
unique_agregados <- length(unique(fumigations$unidade))
people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
people_observed <- length(unique(model_data$oracle_number))
pr <- function(x){prettyNum(x, big.mark = ',')}
```

**Fumigations**: During the period from January 1st, 2013 through December 31st, 2016, the Maragra Malaria Control Unit carried out `r pr(n_fumigations)` episodes of fumigation of residential "agregados" (household combinations), for a total of `r pr(n_buildings)` building-fumigation combinations. The total number of unique agregados sprayed during this period was `r pr(unique_agregados)`. Among the `r pr(people_observed)` workers for whom we have reliable absenteeism and residential data, `r pr(people_sprayed)` had their homes fumigated at least once (the majority of workers live off of the facility).

```{r}
dry_days <- length(which(weather$precipitation > 0))
n_days <- length(weather$precipitation)
p_dry <- round(dry_days / n_days * 100)
amount_rain <- round(mean(weather$precipitation[weather$precipitation > 0], na.rm = T), digits = 2)
x <- weather %>%
  group_by(month = format(date, '%m')) %>%
  summarise(precipitation = mean(precipitation, na.rm = TRUE))
```



**Absences**: We observed `r pr(nrow(ab_panel))` unique worker-days among the `r pr(people_observed)` workers. The all-period average absenteeism rate was `r round(100 * sum(ab_panel$absent) / nrow(ab_panel), digits = 2)`%, though this rate varied widely as a function of worker department, sex, residence, and season (table 1).


```{r, results='asis'}
calculate <- function(var){
  out <- 
  model_data %>%
    mutate(year = format(date, '%Y')) %>%
    mutate(on_site = ifelse(on_site,
                                    'On site',
                                    'Off site'),
           rainy = ifelse(rainy,
                          'Rainy',
                          'Dry')) %>%
    filter(!is.na(rainy)) %>%
  group_by_(var, 'year') %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent),
            workers = length(unique(oracle_number))) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) 
  names(out)[1] <- 'variable'
  out$val <- paste0(round(out$absenteeism_rate, digits = 1),
                    '%')#,
                    # ')',
                    # '\n',
                    # ' (workers:', out$workers,
                    # ')')
  out <- 
    out %>%
    dplyr::select(variable, year, val)
  
  out <- out %>%
    spread(key = year,
           value = val)
  out$variable <- as.character(out$variable)
  new_var <- ifelse(var == 'on_site',
                    'Residence',
                    ifelse(var == 'rainy',
                           'Precipitation',
                           var))
  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
  out <- cbind(left, out)
  names(out)[1:2] <- c('Variable', 'x')
  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
  return(out)
}
out_list <- list()
vars <- c('season', 
           'field',
          'permanent_or_temporary',
           'sex',
           'on_site',
           'rainy')
for(i in 1:length(vars)){
  x <- calculate(var = vars[i])
  out_list[[i]] <- x
}
ab_table <- bind_rows(out_list)
ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
ab_table$x <- Hmisc::capitalize(ab_table$x)
names(ab_table)[2] <- ''
kable(ab_table, format = 'html') %>% kable_styling()
# print(xtable::xtable(ab_table,
#                      caption = 'Absenteeism rate by year and worker characteristics',
#                      align = c('r', 'r', 'l', rep('l', 4))),
#       include.rownames = FALSE)
```


```{r}
ov_ab <- round(100 * mean(ab_panel$absent[between(ab_panel$date, as.Date('2012-01-01'), as.Date('2016-12-31'))]), digits = 2)

ov_cl <- length(which(between(clinic$date, as.Date('2012-01-01'), as.Date('2016-12-31'))))
```

**Precipitation**: Of the `r n_days` days observed, `r dry_days` had no rainfall at all (ie, approximately two-thirds). On days for which there was any rainfall at all, the average amount was approximately `r amount_rain` centimeters. Rain was most common in December and January (average of 5-5.5 cm daily) and least common in August and September (average of 0.5 cm daily). Days with any rainfall saw more absenteeism than days with no rainfall (Figure 6, panel A) and among days with any rainfall, more precipitation was associated with greater absenteeism (Figure 6, panel B) (correlation coefficient of 0.25).

```{r, fig.cap = 'Rainfall and absenteeism: association of any rainfall with absenteeism rate (left) and association of rainfall amount and absenteeism (right)', fig.height=3.5, fig.width=7, fig.align='center'}
bi_data <- model_data %>%
  group_by(date) %>%
  mutate(any_rain = mean(precipitation, na.rm = TRUE) > 0) %>%
  ungroup %>%
  filter(!is.na(any_rain)) %>%
  mutate(any_rain = ifelse(any_rain, 'Some rain', 'No rain')) %>%
  group_by(any_rain, date) %>%
  summarise(absenteeism = mean(absent) * 100)
bi_data_simple <- bi_data %>%
  group_by(any_rain) %>%
  summarise(absenteeism = mean(absenteeism))
p1 <- ggplot(data = bi_data,
       aes(x = any_rain,
           y = absenteeism)) +
  geom_jitter(alpha = 0.3, size = 0.5) +
  geom_violin(alpha = 0.4, fill = 'lightblue') +
  scale_y_log10() +
  theme_maragra() +
  geom_point(data = bi_data_simple,
             aes(x = any_rain,
                 y = absenteeism),
             alpha = 0.8) +
  geom_text(data = bi_data_simple,
             aes(x = any_rain,
                 y = absenteeism,
                 label = paste0(round(absenteeism, digits = 2), '\n(median)')),
             color = 'black') +
  labs(x = '',
       y = 'Absenteeism rate (log scale)',
       title = 'A')

plot_data <- model_data %>%
  group_by(date) %>%
  summarise(precipitation = mean(precipitation, na.rm = TRUE),
            absenteeism = mean(absent)) %>%
  ungroup %>%
  filter(precipitation > 0.1)
p2 <- ggplot(data = plot_data,
       aes(x = precipitation,
           y = absenteeism * 100)) +
  geom_point(alpha = 0.3, size = 0.5) +
  scale_x_log10() +
  geom_smooth() +
  # scale_y_log10() +
  theme_maragra() +
  labs(x = 'Daily precipitation (log scale)',
       y = 'Absenteeism rate', 
       title = 'B')
Rmisc::multiplot(p1, p2, cols = 2)
```




```{r, fig.align = 'center', fig.cap = 'Distribution of herd protection score in panel data'}
ggplot(data = model_data,
       aes(x = herd)) +
  geom_density(fill = 'lightblue',
               alpha = 0.6) +
  theme_maragra() +
  labs(x = 'Herd protection score',
       y = 'Density')
```

\newpage


```{r, results = 'asis'}
make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results")
```


## Sensitivity analysis

The below shows average FE for year / worker, during malaria season, on a rainy day...

```{r}
expander <- function(x){sort(unique(unlist(model_data[,x])))}
dummy <- expand.grid(season = expander('season'),
                     months_since = expander('months_since'),
                     rainy_day = expander('rainy_day'),
                     herd = mean(model_data$herd),
                     group = expander('group'))
dummy$predicted <- NA

groups <- sort(unique(dummy$group))
for (i in 1:length(groups)){
    message(i)
    this_group <- groups[i]
    indices <- which(dummy$group == this_group)
    model <- protection_models[[this_group]]
    fe <- getfe(model)
    # Add back in the mean effect of fixed effects
    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
    data <- dummy[indices,]  
    # predict_felm(model = model, data = )
    dummy$predicted[indices] <- 
      (regtools::predict.felm(object = model,
                             newdata = data,
                             use.fe = FALSE) + add) * 100
}

ggplot(data = dummy %>% filter(rainy_day),
       aes(x = months_since,
           y = predicted,
           color = season,
           group = season)) +
  geom_line() +
  geom_point() +
  facet_wrap(~group, scales = 'free_y') +
  scale_colour_manual(name = 'Malaria\nseason',
                      values = c('blue', 'red')) +
  theme_maragra() +
  labs(x = 'IRS status',
       y = 'Estimated absenteeism') #+
  # theme(legend.position = 'right')


```

The below shows estimated importance of herd protection, and heterogeniety across groups.

```{r}
dummy <- expand.grid(season = expander('season'),
                     months_since = expander('months_since'),
                     rainy_day = expander('rainy_day'),
                     herd = 0:1000,
                     group = expander('group'))
dummy$predicted <- NA


groups <- sort(unique(dummy$group))
for (i in 1:length(groups)){
    message(i)
    this_group <- groups[i]
    indices <- which(dummy$group == this_group)
    model <- protection_models[[this_group]]
    fe <- getfe(model)
    # Add back in the mean effect of fixed effects
    add <- fe %>% group_by(fe) %>% summarise(effect = mean(effect)) %>% summarise(sum(effect)) %>% unlist %>% as.numeric
    data <- dummy[indices,]  
    dummy$predicted[indices] <- 
      (regtools::predict.felm(object = model,
                             newdata = data,
                             use.fe = FALSE) + add) * 100
}

ggplot(data = dummy %>% filter(rainy_day, season == 'high'),
       aes(x = herd,
           y = predicted,
           color = months_since,
           group = months_since)) +
  geom_line() +
  facet_wrap(~group, scales = 'free_y') +
  scale_colour_manual(name = 'IRS status',
                      values = c('darkorange', 'darkgreen')) +
  theme_maragra() +
  labs(x = 'Herd protection score',
       y = 'Estimated absenteeism')
```


### What would happen without IRS?

The below chart shows the modeled average absenteeism rate during malaria season for permanent workers, by group type.

```{r}
x <- model_data %>%
  dplyr::filter(season == 'high'#,
                # group == groups[1],
                # permanent_or_temporary == 'Permanent'
                )
x <- x [,grepl('predicted', names(x)) | names(x) == 'group']
x <- x %>% gather(key, value, predicted_max_herd_max_irs:predicted)
dict <- data.frame(key = c('predicted', 
                           'predicted_no_herd',
                           'predicted_no_irs',
                           'predicted_no_herd_no_irs',
                           'predicted_max_herd_max_irs',
                           'predicted_max_herd',
                           'predicted_max_irs'),
                   new_key = c('As is',
                               'No herd protection',
                               'No individual protection',
                               'No protection (herd or individual)',
                               'Protection: maximum IRS and herd',
                               'Protection: maximum herd',
                               'Protection: maximum IRS'))

x <- left_join(x,
               dict)
plot_data <- x %>%
  mutate(new_key = gsub(' ', '\n', new_key)) %>%
  group_by(group, 
           new_key) %>%
  summarise(avg = mean(value, na.rm = TRUE),
            med = median(value, na.rm = TRUE),
            upr = quantile(value, 0.75, na.rm = TRUE),
            lwr = quantile(value, 0.25, na.rm = TRUE))
levs <- gsub(' ', '\n', c('No herd protection',
                               'No individual protection',
                               'No protection (herd or individual)',
                               'As is',
                               'Protection: maximum IRS',
                               'Protection: maximum herd',
                               'Protection: maximum IRS and herd'))
plot_data$new_key <- factor(plot_data$new_key,
                       levels = levs)
cols <- colorRampPalette(brewer.pal(n = 9,
                                    name = 'Set1'))(length(unique(plot_data$group)))

library(ggrepel)
ggplot(data = plot_data,
       aes(x = new_key,
           y = avg * 100,
           color = group)) +
  geom_point(size = 4) +
  theme_maragra() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = 'Scenario',
       y = 'Absenteeism rate') +
  geom_line(aes(group = group),
            alpha = 0.5,
            size = 1) +
  scale_color_manual(name = '',
                     values = cols) #+
  # ggrepel::geom_label_repel(aes(label = round(avg * 100, digits = 1)))
```

## Return on investment

- Number of working days

```{r, echo = TRUE}
x <- ab_panel %>% group_by(date) %>% tally
ggplot(data = x,aes(x = date,y = n)) + geom_line() +
  labs(title = mean(x$n))
```

- Number of working days annually 

```{r, echo = TRUE}
options(scipen = '999')
x <- ab_panel %>% mutate(year = as.numeric(format(date, '%Y'))) %>% group_by(year) %>% tally
ggplot(data = x,aes(x = year,y = n)) + geom_line() +
  labs(title = mean(x$n))
```

- Number of prevented absences per year

```{r, echo = TRUE}
ratio <- length(unique(ab_panel$oracle_number)) / length(unique(model_data$oracle_number))

x <- model_data %>%
  group_by(year = calendar_year) %>%
  summarise(eligibles = n(),
            absences = mean(as.numeric(absent), na.rm = TRUE) * n(),
            counterfactual = mean(predicted_no_herd_no_irs,na.rm =TRUE) * n()) 
y <- x %>%
  summarise(x = mean(counterfactual) - mean(absences)) %>%
  unlist * ratio
# Annual prevented
y
x <- x %>%
  mutate(prevented = counterfactual - absences)
# Number of presences to prevent 1 case
sum(x$eligibles) / sum(x$prevented)
```

```{r, eval = TRUE}
# Clinical costs
(y /
  80) * # number of absences it takes to produce one company clinic positve
  31.49 # cost per clinical case per ezenduka

# Wage costs
y *
  4 # assuming USD
```

\newpage

## Robustness and generalizability 

```{r, echo = FALSE}
ro_data <- model_data %>%
# ro_data <- ab_panel %>%
#   left_join(irs, by = c('date', 'unidade')) %>%
#   filter(days_since < 0) %>%
  mutate(pre_irs = between(days_since, -10, -1)) 
#   # mutate(time_period = lendable::time_period_extract(date)) %>%
#   # mutate(time_period = as.character(time_period)) %>%
#   mutate(time_period = make_season(date = date)) %>%
#   mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) %>%
#   left_join(workers %>%
#               dplyr::select(oracle_number,
#                             sex,
#                             department,
#                             permanent_or_temporary),
#             by = 'oracle_number')
# ro_data$time_period <- as.character(ro_data$time_period)


make_model <- function(formula = 'absent ~ pre_irs + time_period'){
  ro_model <- glm(as.formula(formula), 
                data = ro_data,
                family = binomial('logit'))
  coefs <- exp(coef(ro_model))
  ci <- confint(ro_model)
  ci <- exp(ci)
  out <- data.frame(Variable = names(coefs),
                    Estimate = coefs,
                    Lower = ci[,1],
                    Upper = ci[,2])
  row.names(out) <- NULL
  out$`P value` <- coef(summary(ro_model))[,4]
  out$Significant <- ifelse(out$`P value` <= 0.001, 'xxx',
                            ifelse(out$`P value` <= 0.01, 'xx',
                                   ifelse(out$`P value` <= 0.05, 'x',
                                          '')))
  out$Variable <- as.character(out$Variable)
  out$Variable <- gsub('pre_irsTRUE', '10 days prior to IRS', out$Variable)
  out$Variable <- gsub('time_period', '', out$Variable)
  return(kable(out, format = "html") %>%
  kable_styling()# %>%
    # row_spec(row = which(grepl('prior', out$Variable)),background = "yellow")
  )
}
```

```{r, echo = FALSE, results = 'asis'}

# # BASIC FORMULA
# make_model('absent ~ pre_irs + time_period')

# # BASIC FORMULA WITH INTERACTIONS
# make_model('absent ~ pre_irs*time_period')

# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX
# make_model('absent ~ pre_irs*time_period + sex')
# 
# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT
# make_model('absent ~ pre_irs*time_period + sex + department')
# 
# 
# # BASIC FORMULA WITH INTERACTIONS AND ADJUSTING FOR SEX AND DEPARTMENT AND STATUS
# make_model('absent ~ pre_irs*time_period + sex + department + permanent_or_temporary')

# Adjust for seasonality and department
make_model('absent ~ pre_irs*season + group + rainy_day')
```



```{r}
# The below shows our robustness check for sick only absenteeism. Unlike with all cause absenteeism, during malaria season, being in the period 10 days prior to IRS is associated with statistically greater likelihood of being absent for illness. In other words, it appears that there is somewhat of a feedback loop: when a worker misses work due to illness, his/her likelihood of getting IRS doubles in the next 10 days.\todo{What to do about this???}
# make_model('absent_sick ~ pre_irs*time_period + department')
```
