---
title: "Final analysis for paper"
date: ''
fig_height: 2.6
output:
  pdf_document:
    includes:
      in_header: preamble.sty
    keep_tex: yes
    latex_engine: xelatex
  html_document:
    df_print: paged
fig_width: 4
always_allow_html: true
subtitle: Brew, Pradhan, Sicuri
bibliography: bibliography.bib
---


```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
library(sp)
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
# library(lme4)
library(lfe)
library(restorepoint)
library(regtools)

loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               # dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
knitr::opts_chunk$set(error = FALSE)

opts_knit$set(width=75)
options(xtable.comment = FALSE)
source('code.R')
```



*Absences*: We observed `r nrow(ab_panel)` unique worker-days among `r length(unique(ab_panel$oracle_number))` workers. The all-period average absenteeism rate was `r round(length(which(ab_panel$absent)) / nrow(ab_panel) * 100, digits = 2)`%, though this rate varied widely as a function of worker department, sex, residence, and season (table).

# Model results (one model)

```{r}
library(broom)
broom::tidy(final_model)

# Prediction based on individual stuff
fake <- expand.grid(protection_var = seq(0, 0.7, by = 0.1),
                                         rain_var = seq(0, 6, 1))
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, newdata = fake, interval = 'confidence'))-1
predictions <- data.frame(predictions)
fixed_effects <- mean(fe$effect)
fake$fit <- predictions$fit + fixed_effects
fake$lwr <- predictions$lwr + fixed_effects
fake$upr <- predictions$upr + fixed_effects

n_cols <- length(unique(fake$protection))
cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, 'Spectral'))(n_cols)
cols[2] <- 'black'
library(maragra)
ggplot(data = fake %>% filter(rain_var == 3) %>% mutate(protection_var = factor(protection_var)),
       aes(x = protection_var,
           y = fit)) +
  geom_ribbon(aes(ymin = lwr, ymax = upr),
              color = NA,
              alpha = 0.15) +
  geom_point() +
  geom_line(alpha = 0.6) +
  labs(x = 'Protection score',
       title = 'Association between community protection, individual protection, and absenteeism',
       y = 'Estimated monthly absenteeism rate\n(Average fixed effect, 3 cm of 15-60 day prior precipitation)') + theme_maragra() 
```

# Model results (segregated models)

```{r}
out_list <- list()
worker_groups <- names(model_list)
for(i in 1:length(worker_groups)){
  this_group <- worker_groups[i]
  out_list[[i]] <- broom::tidy(model_list[[i]]) %>% mutate(grp = this_group)
}
out <- bind_rows(out_list)
# Get the non segrated version too
non_seg <- broom::tidy(final_model) %>% mutate(grp = 'Non segregated')
options(scipen =  '999')
out <- out %>% bind_rows(non_seg) %>% dplyr::select(term, estimate, p.value, grp) %>%
  mutate(Type = 'Distance weighted')
out

out %>% mutate(estimate = paste0(round(estimate, digits = 3),
                                 ' (',
               round(p.value, digits = 3), ')')) %>% dplyr::select(-p.value, -Type) %>%  tidyr::spread(key = grp, value = estimate)
```


```{r}
# Also make predictions by groups
worker_groups <- names(model_list)
the_groups <- c(worker_groups, 'All')

# Do predictions on the different groups
fake <- expand.grid(protection_var = seq(0, 1, by = 0.1),
                    rain_var = mean(final_data$rain_var, na.rm = TRUE),
                    group = the_groups)
fake_list <- list()
for(i in 1:length(the_groups)){
  this_worker_group <- the_groups[i]
  if(this_worker_group == 'All'){
    this_model <- final_model
    this_model_lm <- final_model_lm
    this_fake_data <- fake %>% mutate(group = 'All')
  } else {
      this_model <- model_list[[i]]
    this_model_lm <- model_list_lm[[i]]
    this_fake_data <- fake %>% dplyr::filter(group == this_worker_group)
  }

  predictions <- exp(predict(this_model_lm, newdata = this_fake_data, interval = 'confidence'))-1
  predictions <- data.frame(predictions)
  this_fe <- getfe(this_model)
  fixed_effects <- mean(this_fe$effect)
  out <- this_fake_data
  out$fit <- predictions$fit + fixed_effects
  out$lwr <- predictions$lwr + fixed_effects
  out$upr <- predictions$upr + fixed_effects
  fake_list[[i]] <- out
}
fake <- bind_rows(fake_list)

n_cols <- length(unique(fake$protection))
cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, 'Spectral'))(n_cols)
cols[2] <- 'black'
library(maragra)
ggplot(data = fake  %>% mutate(protection = factor(protection_var)),
       aes(x = protection_var,
           y = fit)) +
  # geom_ribbon(aes(ymin = lwr, ymax = upr,
  #                 fill = protection),
  #             color = NA,
  #             alpha = 0.15) +
  geom_point() +
  facet_wrap(~group) +
  geom_line(alpha = 0.6) +
  labs(x = 'Protection score',
       title = 'Association between protection score and absenteeism',
       y = 'Estimated monthly absenteeism rate') +    
  theme_maragra() 

```


# Simulations

```{r}

# SIMULATIONS
sim_data <- simulations
sim_data$idx <- sim_data$oracle_number
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
predictions <- data.frame(predictions)
# Add the fixed effects
fixed_effects <- left_join(sim_data,
                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
                             mutate(idx = as.character(idx))) %>%
  .$effect
fixed_effects <- as.numeric(fixed_effects)

predictions$fit <- predictions$fit + fixed_effects
predictions$lwr <- predictions$lwr + fixed_effects
predictions$upr <- predictions$upr + fixed_effects

# combine other info
predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
  mutate(absent = as.numeric(absent)) %>%
  filter(!is.na(fit))

simple <- predictions %>%
  group_by(strategy) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE))# %>%
  # dplyr::rename(value = predicted) %>%
  # dplyr::select(strategy, value, predicted_absences)

zero <- predictions %>%
  filter(strategy == 'Zero') %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  group_by(year) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences)
# dplyr::rename(value = predicted) %>%
# dplyr::select(permanent_or_temporary, value, predicted_absences)

simple
# kable(simple)

# Eligible working days
sum(zero$n)
# Absences
sum(zero$actual_absences)
# Actual absence rate
sum(zero$actual_absences) / sum(zero$n)
# Would have observed
sum(zero$predicted_absences)
# Predicted - observ
sum(zero$predicted_absences) - sum(zero$actual_absences)
  # Overall reduction from
sum(zero$predicted_absences) / sum(zero$n)
# To
sum(zero$actual_absences) / sum(zero$n)
# Number of working days permanent vs temp
table(predictions$permanent_or_temporary[predictions$strategy=='Real'])
# Absences avoided among permanent workers
x = predictions %>% filter(strategy == 'Zero') %>% group_by(permanent_or_temporary) %>% summarise(n = n(),
                                                                                              actual_absences = sum(absent),
                                                                                              predicted_absences = sum(fit, na.rm = TRUE),
                                                                                              actual = mean(absent, na.rm = TRUE),
                                                                                              predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences) %>%
  # decline due to irs
  mutate(decline = predicted - actual)
x
# Weight temporary twice
weighted <- x$
# One absence every n days
sum(zero$n) / sum(zero$avoided)
# Annual reduction in absence
mean(zero$avoided[zero$year < 2016])
# Cost per absence avoided
68984 / 4183
# Percentage of absence among permanent workers
x = predictions %>%
  filter(strategy == 'Zero') %>%
  # filter(date < '2016-01-01') %>%
  group_by(permanent_or_temporary) %>%
  summarise(absences = sum(absent),
            predicted = sum(fit)) %>%
  mutate(averted = predicted - absences) %>%
  mutate(p = averted / sum(averted))
x
# Weighted average cost
y = sum(c(25, 5) * x$p)
y
# Cost of treatment per absence
22 / 80
# Total savings per aversion
0.275 + y
# Total yearly savings
(0.275 + y) * 4183
# Total yearly savings
((0.275 + y) * 4183) - 68984
# ROI
((0.275 + y) * 4183) / 68984
# Additional max strategy
simple
47943-42131

ggplot(data = simple,
       aes(x = strategy,
           y = value)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(value, digits = 3)),
             nudge_y = -0.005)
```

# Descriptive statistics

```{r}
table(workers$sex)
x = model_data %>% dplyr::distinct(oracle_number, .keep_all=T); prop.table(table(x$sex))
prop.table(table(x$field))
x = left_join(x, workers %>% dplyr::select(oracle_number, date_of_birth)) %>% mutate(age = as.numeric(date - date_of_birth) / 365.25)
hist(x$age)
prop.table(table(x$age >= 20 & x$age < 40))
prop.table(table(x$permanent_or_temporary))
```


# Table 1

```{r}
library(table1)
message('We observed ', nrow(model_data), ' worker-days among ', length(unique(model_data$oracle_number)), ' unique workers. Overall absence rate was ', length(which(model_data$absent)) / nrow(model_data) * 100)

# Table of those absences
vars <- list('field',
          'permanent_or_temporary',
           'sex')
table1::label(model_data$field) <- 'Worker location'
table1::label(model_data$permanent_or_temporary) <- 'Worker contract'
table1::label(model_data$sex) <- 'Worker sex'
model_data$year <- as.numeric(format(model_data$date, '%Y'))
table1::label(model_data$year) <- 'Year'

table1::table1(~field + permanent_or_temporary + sex | year , data = model_data)
```

# Crude effect on absences chart

```{r}
pd <- model_data %>%
  mutate(days_since = round(days_since / 30)) %>%
  group_by(days_since) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(p = absences / eligibles * 100) %>%
  filter(days_since >= -10,
         days_since <= 10)

g1 <- ggplot(data = pd,
       aes(x = days_since,
           y = p)) +
  geom_step(color = 'darkred', alpha = 0.8, size = 1) +
  theme_maragra() +
  labs(x = 'Months relative to administration of IRS',
       y = 'Absenteeism rate (%)',
       title = 'A') +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.7)
pd2 <- model_data %>%
  mutate(days_since = round(days_since / 30)) %>%
  mutate(season = ifelse(rain_var >= median(rain_var, na.rm = TRUE),
                         'Rainy season',
                         'Dry season')) %>%
  group_by(days_since, season) %>%
  summarise(absences = length(which(absent)),
            eligibles = n()) %>%
  ungroup %>%
  mutate(p = absences / eligibles * 100) %>%
  filter(days_since >= -10,
         days_since <= 10,
         !is.na(season))
g2 <- ggplot(data = pd2,
       aes(x = days_since,
           y = p)) +
  facet_wrap(~season, ncol = 1) +
  geom_step(color = 'darkred', alpha = 0.8, size = 1) +
  theme_maragra() +
  labs(x = 'Months relative to administration of IRS',
       y = 'Absenteeism rate (%)',
       title = 'B') +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.7)

Rmisc::multiplot(g1, g2, cols = 2)
```


# Absences


