---
title: "Concept note: additional models"
subtitle: "Pradhan, Sicuri, Brew"
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---


```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)

```

```{r, eval = TRUE}
source('paper_prepare.R')
```


\begin{center}
\begin{large}

A few different approaches to modeling absenteeism in the Maragra data

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Overview}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm} 

This document contains methods and basic results for a few different modeling approaches, following a conversation and write-up with Menno Pradhan in May 2018. The purpose of this document is to make explicit the current models we have implemented, and some issues with them.

\end{changemargin}
\end{center}

\vspace{20mm}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Main points}
        \begin{itemize}
          \item I have implemented Menno's suggestions for more simple models.
          \item The simplest models show no overall effect from IRS when not accounting for malaria season
          \item I believe that the lack of an effect is due to seasonality's presence in the model as a fixed (additive) effect, rather than as an interaction with IRS
        \end{itemize}
        \vspace{2mm}
    }%
}

\vfill
\null

\subsection*{Desinataires}
\textbf{Elisa Sicuri; Menno Pradhan}

\vspace{3mm}

\newpage

# The current approach 

In the current draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}
), the model formula is:


$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * IRS_{it} + \hat{\beta}_2{RainyDay_{t}} + \hat{\beta}_3{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
$$

In R, using the `felm` library, this is written as follows:

```{r, echo = FALSE}
protection_models[[1]]$call
```

The above notation shows `absent` (binary 1,0) as a function of the interaction of binary `season` (high vs. low) and binary `months_since` (before vs. after) plus binary `rainy_day` (whether or not it rained on the particular day) plus continuous numeric `herd` (ie, the protection afforded by others) plus fixed effects for both the worker (`oracle_number`) and the year (`malaria_year`).  $\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season. $\upsilon$ is the error term. This model is run separately for the 4 different worker groups, yielding the following results:

```{r, results = 'asis'}
make_models_table(model_list = protection_models, the_caption = "All absenteeism with herd immunity: model results", type = 'latex', multiplier = 1)
```

In the above, "after" IRS application (ie, the 6 months following IRS) is associated with a significant reduction in absenteeism for both kinds of permanent workers, and does not have a signficiant association with a change in absenteeism for both kinds of temporary workers.

# Suggested changes from Menno


Menno sent a write-up on May 17th (see email with subject "suggestions for models") with some proposed changes to how we model absenteeism. What follows is an implementation of those suggestions. 

## Model A (without seasonality)

The below is a characterization of the approach Menno suggested:

- We use monthly instead of daily data (modeling a percentage absenteeism rate rather than a binary).
- We use fixed effects for year-month combination. 
- We don't explicitly model for malaria season. 
- We don't have any interaction terms.
- We use dummies for months since IRS.
- We use the most recent spraying as the time since IRS (ie, if 5 months ago, and 1 month ago, then the dummy is applied for 1 month ago but not 5 months ago).


Using this approach, our panel data looks like this:

```{r}
# Create a model per menno's specifications
menno_models <- list()

menno_data <- model_data %>%
  # filter(!is.na(days_since)) %>%
  # filter(days_since >= -184, days_since <= 184) %>%
  group_by(oracle_number, group, year_month) %>%
  summarise(absent = length(which(absent)) / length(absent) * 100,
            days = n(),
            herd = mean(herd, na.rm = TRUE),
            months_since_menno = dplyr::first(months_since_menno),
            season = dplyr::first(season)) %>%
  ungroup
groups <- sort(unique(model_data$group))
# library(lmerTest)
# library(nlme)
for (i in 1:length(groups)){
  message(i)
  this_group <- groups[i]
  message(this_group)
  these_data <- menno_data %>% filter(group == this_group)
  this_menno_model <- felm(absent ~ months_since_menno    | oracle_number + year_month | 0 | 0,
                                data = these_data)

  menno_models[[i]] <- this_menno_model
}
names(menno_models) <- groups
print(head(menno_data %>% filter(months_since_menno != 'Before') %>% sample_n(20) %>% dplyr::select(-season, -herd)))
```

The model formula was described in Menno's email.

\includegraphics{images/menno}[width=3cm]

Which translates into R as:

```{r}
menno_models[[1]]$call
```

Note that the `months_since_menno` variable has the following categories, and is treated as a "wide" dummy by the software in modeling:

```{r}
levels(menno_data$months_since_menno)
```

The results of this model look like this:


```{r, results = 'asis'}
make_models_table(model_list = menno_models, the_caption = "Simple models with year-month fixed effects", type = 'latex', multiplier = 1)
```

In the above, it is apparent that the effects of IRS appear to be either (a) in the opposite direction as expected or (b) non-signficant. Here is a visualization of IRS' apparent effect (relative to the "Before" period, which is the reference class). Were IRS effective, we would expect the black line to be below 0 (the red line).

```{r}
out_list <- list()
for (i in 1:length(groups)){
  this_group <- groups[[i]]
  x <- coef(menno_models[[i]])
  x <- x[grepl('months_since', names(x))]
  names(x) <- 1:6
  out_list[[i]] <- data.frame(group = this_group,
                              month = names(x),
                              coef = x)
}
out <- bind_rows(out_list)

ggplot(data = out,
       aes(x = month,
           y = coef)) +
  geom_line(aes(group = 1)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0, 
             lty = 2,
             col = 'red',
             alpha = 0.6) +
  labs(x = 'Months since IRS',
       y = 'Coefficient')
```

What is going on here? Could it  be that IRS has a positive (ie, increases) impact on absenteeism? Let's look at the raw data, rather than the coefficients. The below chart shows absenteeism (as a percentage of the pre-IRS absenteeism average) for workers, broken down by season (row) and worker type (column), with point size reflecting the number of observations.

```{r, fig.width = 8}
pd <- menno_data %>%
  group_by(season, months_since_menno, group) %>%
  summarise(absenteeism_rate = mean(absent, na.rm = TRUE),
            days = sum(days)) %>%
  # summarise(absenteeism_rate = length(which(absent)) / length(absent) * 100) %>%
  ungroup %>%
  group_by(group, season) %>%
  mutate(start = dplyr::first(absenteeism_rate)) %>%
  mutate(p = absenteeism_rate / start * 100)

options(scipen = '999')
ggplot(data = pd %>% ungroup %>% 
         mutate(season = paste0(season, ' season')) %>%
         mutate(group = gsub(' worker', '', group)),
       aes(x = months_since_menno,
           y = p)) +
  geom_point(aes(size = days)) +
  geom_line(aes(group = 1), alpha = 0.4) +
  facet_grid(season~group) +
  geom_hline(yintercept = 100, 
             color = 'red',
             lty = 2) +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_log10(breaks = seq(0, 500, 100)) +
  labs(y = 'Absenteeism (relative to before IRS, log)',
       x = 'Time') +
  scale_size_continuous(name = 'Observations',
                    breaks = c(10000, 100000, 5000000))

```

In the above chart, two things stand out:

1. Variability is greatest among temporary workers.
2. There appears to be a reduction in absenteeism post IRS for permanent workers _during the malaria season_, but not during the non-malaria season

So, perhaps controlling for year-month through fixed effects isn't enough, since the effect of IRS _interacts_ with the season...


## Model B (with seasonality)

Using this approach, our panel data looks like this:

```{r}
# Create a model per menno's specifications
menno_models <- list()

for (i in 1:length(groups)){
  message(i)
  this_group <- groups[i]
  message(this_group)
  these_data <- menno_data %>% filter(group == this_group)
  this_menno_model <- felm(absent ~ months_since_menno*season  | oracle_number + year_month | 0 | 0,
                                data = these_data)

  menno_models[[i]] <- this_menno_model
}
names(menno_models) <- groups
print(head(menno_data %>% filter(months_since_menno != 'Before') %>% sample_n(20) %>% dplyr::select(-herd)))
```

Our new model specification (in R) is:

```{r}
menno_models[[1]]$call
```

The results of this model look like this:


```{r, results = 'asis'}
make_models_table(model_list = menno_models, the_caption = "Simple models with year-month fixed effects and seasonality-IRS interaction term", type = 'latex')
```

The below shows model coefficients:

```{r, fig.width = 8}
out_list <- list()
for (i in 1:length(groups)){
  this_group <- groups[[i]]
  x <- clean_up_model(menno_models[[i]], multiplier = 1)
  x <- x %>%
    filter(grepl('IRS status', Term),
           grepl('season', Term)) 
  # Malaria season
  df1 <- data.frame(group = this_group,
                              month = 1:6,
                              coef = x$Estimate,
                    season = 'high')
  
  x <- clean_up_model(menno_models[[i]], multiplier = 1)
  x <- x %>%
    filter(grepl('IRS status', Term),
           !grepl('season', Term)) 
  # Malaria season
  df2 <- data.frame(group = this_group,
                              month = 1:6,
                              coef = x$Estimate,
                    season = 'low')
  out_list[[i]] <- bind_rows(df1, df2) 
}
out <- bind_rows(out_list)
x <- strsplit(out$coef, ' ')
x <- lapply(x, function(x){x[1]})
x <- unlist(x)
out$coef <- as.numeric(as.character(x))
ggplot(data = out,
       aes(x = month,
           y = coef)) +
  geom_line(aes(group = 1)) +
  facet_grid(season~group) +
  geom_hline(yintercept = 0, 
             lty = 2,
             col = 'red',
             alpha = 0.6) +
  labs(x = 'Months since IRS',
       y = 'Coefficient')
```


## Model C (with seasonality and herd score)

We now introduce herd immunity to the model. This is introduced as an additive effect; perhaps it should be interactive with IRS? Using this approach, our panel data looks like this:

```{r}
# Create a model per menno's specifications
menno_models <- list()

for (i in 1:length(groups)){
  message(i)
  this_group <- groups[i]
  message(this_group)
  these_data <- menno_data %>% filter(group == this_group)
  this_menno_model <- felm(absent ~ months_since_menno*season  | oracle_number + year_month | 0 | 0,
                                data = these_data)

  menno_models[[i]] <- this_menno_model
}
names(menno_models) <- groups
print(head(menno_data %>% filter(months_since_menno != 'Before') %>% sample_n(20) ))
```

Our new model specification (in R) is:

```{r}
menno_models[[1]]$call
```

The results of this model look like this:


```{r, results = 'asis'}
make_models_table(model_list = menno_models, the_caption = "Simple models with year-month fixed effects, seasonality-IRS interaction term, and herd score", type = 'latex')
```

